---
title: "Data Analysis"
author: "Stella Lee, Yoon Choi, Aru Fatehpuria"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Packages
library(dplyr)
library(ggplot2)
library(mosaic)
library(stringr)
library(leaps)
library(gridExtra)
library(cowplot)
```

# Upload Dataset

```{r warning=FALSE}
HPIdata <- read.csv("https://raw.githubusercontent.com/stellasylee/Happiness-Planet-Index/master/finalDataset.csv")[-1]
# Change columns into numeric
HPIdata$RegionNum <- as.numeric(substring(HPIdata$Region, 1, 1))
HPIdata$GovHealth <- as.numeric(as.character(HPIdata$GovHealth))
HPIdata$PriHealth <- as.numeric(as.character(HPIdata$PriHealth))
HPIdata$PPP <- as.numeric(as.character(HPIdata$PPP))
HPIdata$PopDensity <- as.numeric(as.character(HPIdata$PopDensity))
HPIdata$WomenParliamant <- as.numeric(as.character(HPIdata$WomenParliamant))
```

Count how many missing rows within a column

```{r}
sum(is.na(HPIdata$SocialSupport))
sum(is.na(HPIdata$UnEmployment))
```

## Main Effect plots without data transformation
```{r warning=FALSE}
Internet <- ggplot(data = HPIdata, aes(x = Internet, y = HPI)) + geom_point()  +
  aes(colour = factor(RegionNum)) + theme(legend.position = "right") + labs(title = "Internet")
LifeExpect <- ggplot(data = HPIdata, aes(x = LifeExpect, y = HPI)) + geom_point()  +
  aes(colour = factor(RegionNum)) + theme(legend.position = "right") + labs(title = "Life Expectency (yr)")
GovHealth <- ggplot(data = HPIdata, aes(x = GovHealth, y = HPI)) + geom_point()  +
  aes(colour = factor(RegionNum)) + theme(legend.position = "right") + labs(title = "Government Health Expenditure")
PriHealth <- ggplot(data = HPIdata, aes(x = PriHealth, y = HPI)) + geom_point()  +
  aes(colour = factor(RegionNum)) + theme(legend.position = "right") + labs(title = "Private Health Expenditure")
PPP <- ggplot(data = HPIdata, aes(x = PPP, y = HPI)) + geom_point()  +
  aes(colour = factor(RegionNum)) + theme(legend.position = "right") + labs(title = "Purchasing Power")
PopDen <- ggplot(data = HPIdata, aes(x = PopDensity, y = HPI)) + geom_point()  +
  aes(colour = factor(RegionNum)) + theme(legend.position = "right") + labs(title = "Population Density")
Unemploy <- ggplot(data = HPIdata, aes(x = UnEmployment, y = HPI)) + geom_point()  +
  aes(colour = factor(RegionNum)) + theme(legend.position = "right") + labs(title = "Unemployment")
Freedom <- ggplot(data = HPIdata, aes(x = FreedomToChoice, y = HPI)) + geom_point()  +
  aes(colour = factor(RegionNum)) + theme(legend.position = "right") + labs(title = "Freedom to Choice")
#grid.arrange(Internet, LifeExpect, GovHealth, PriHealth, PPP, PopDen, Unemploy, Freedom, ncol = 2)
grid.arrange(Internet, LifeExpect, GovHealth, PriHealth, ncol = 2)
grid.arrange(PPP, PopDen, Unemploy, Freedom, ncol = 2)
```

## Transformation in explanatory variables

```PopDensity```, ```PPP```, ```GovHealth```, ```PriHealth``` contains extreme values -> log transformation

```{r}
GovHealth <- ggplot(data = HPIdata, aes(x = GovHealth, y = HPI)) + geom_point()  +
  aes(colour = factor(RegionNum)) + theme(legend.position = "right") + labs(title = "Government Health Expenditure")
GovHealthlog <- ggplot(data = HPIdata, aes(x = log(GovHealth), y = HPI)) + geom_point()  +
  aes(colour = factor(RegionNum)) + theme(legend.position = "right") + labs(title = "Government Health Expenditure")
PriHealth <- ggplot(data = HPIdata, aes(x = PriHealth, y = HPI)) + geom_point()  +
  aes(colour = factor(RegionNum)) + theme(legend.position = "right") + labs(title = "Private Health Expenditure")
PriHealthlog <- ggplot(data = HPIdata, aes(x = log(PriHealth), y = HPI)) + geom_point()  +
  aes(colour = factor(RegionNum)) + theme(legend.position = "right") + labs(title = "Private Health Expenditure")
PPP <- ggplot(data = HPIdata, aes(x = PPP, y = HPI)) + geom_point()  +
  aes(colour = factor(RegionNum)) + theme(legend.position = "right") + labs(title = "Purchasing Power")
PPPlog <- ggplot(data = HPIdata, aes(x = log(PPP), y = HPI)) + geom_point()  +
  aes(colour = factor(RegionNum)) + theme(legend.position = "right") + labs(title = "Purchasing Power")
PopDen <- ggplot(data = HPIdata, aes(x = PopDensity, y = HPI)) + geom_point()  +
  aes(colour = factor(RegionNum)) + theme(legend.position = "right") + labs(title = "Population Density")
PopDenlog <- ggplot(data = HPIdata, aes(x = log(PopDensity), y = HPI)) + geom_point()  +
  aes(colour = factor(RegionNum)) + theme(legend.position = "right") + labs(title = "Population Density")
grid.arrange(GovHealth, GovHealthlog, PriHealth, PriHealthlog, ncol = 2)
grid.arrange(PPP, PPPlog, PopDen, PopDenlog, ncol = 2)
```

## Best Subset

```{r}
# Design matrix for best subset
full_data=cbind(factor(HPIdata$Region), factor(HPIdata$Year), HPIdata$Internet, HPIdata$LifeExpect, HPIdata$GovHealth, log(HPIdata$GovHealth), HPIdata$PriHealth, log(HPIdata$PriHealth), HPIdata$PPP, log(HPIdata$PPP), HPIdata$PopDensity,log(HPIdata$PopDensity),HPIdata$UnEmployment, HPIdata$FreedomToChoice, HPIdata$Corruption)
row.has.na <- apply(HPIdata, 1, function(x){any(is.na(x))})
HPIdata.filtered <- HPIdata[!row.has.na,]
full_explanatory_filtered <- full_data[!row.has.na,]

# Best subset
best.lm <- leaps(full_explanatory_filtered,HPIdata.filtered$HPI, method = "adjr2", names = c('Region','Year','Internet','Life Expectancy','Gov Health','log Gov Health', 'PriHealth','log PriHealth','PPP','log PPP', 'PopDen','Log PopDensity','Unemployment','Freedom to Choice','Corruption'), nbest=3)
data1 = cbind(best.lm$which,best.lm$adjr2)

# 8 variables - R2 of 0.751
model_trans_8 <-  lm (HPI ~ factor(Region) + Internet + LifeExpect + log(GovHealth) + PPP + log(PPP) + log(PopDensity) + FreedomToChoice, data = HPIdata)
anova(model_trans_8)
summary(model_trans_8)

# 9 variable model - R2 of  0.7842 
row.has.na.final <- apply(HPIdata[1:15], 1, function(x){any(is.na(x))})
finalData <- HPIdata[!row.has.na.final,]
model_trans_9 <- lm (HPI ~ factor(Region) + Internet + LifeExpect + log(GovHealth) + PriHealth + log(PriHealth) + PPP +
                         log(PopDensity) + UnEmployment,
                       data = HPIdata)
anova(model_trans_9)
summary(model_trans_9)

# With 10 variables - R2 of 0.7843
model_trans_10 <-  lm (HPI ~ factor(Region) + Internet + LifeExpect + log(GovHealth) + PriHealth + log(PriHealth) + PPP + log(PPP) + log(PopDensity) + UnEmployment, data = HPIdata)
anova(model_trans_10)
summary(model_trans_10)
```

## Final Model

```{r}
row.has.na.final <- apply(HPIdata[1:15], 1, function(x){any(is.na(x))})
finalData <- HPIdata[!row.has.na.final,]

finalData$Region <- as.character(finalData$Region)
finalData$Region[finalData$Region %in% "1a"] <- "Central America, Mexico & Carribean"
finalData$Region[finalData$Region %in% "1b"] <- "South America"
finalData$Region[finalData$Region %in% "2a"] <- "Australia & NZ"
finalData$Region[finalData$Region %in% "2b"] <- "North America"
finalData$Region[finalData$Region %in% "2c"] <- "Western Europe"
finalData$Region[finalData$Region %in% "2d"] <- "Nordic Europe"
finalData$Region[finalData$Region %in% "2e"] <- "Southern Europe"
finalData$Region[finalData$Region %in% "3a"] <- "North Africa"
finalData$Region[finalData$Region %in% "3b"] <- "Middle East / South West Asia"
finalData$Region[finalData$Region %in% "4a"] <- "Southern & Central Africa"
finalData$Region[finalData$Region %in% "4b"] <- "East Africa"
finalData$Region[finalData$Region %in% "4c"] <- "West Africa"
finalData$Region[finalData$Region %in% "5a"] <- "South Asia"
finalData$Region[finalData$Region %in% "6a"] <- "China"
finalData$Region[finalData$Region %in% "6b"] <- "Wealthy East Asia"
finalData$Region[finalData$Region %in% "6c"] <- "South East Asia"
finalData$Region[finalData$Region %in% "7a"] <- "Central Asia & Caucuses"
finalData$Region[finalData$Region %in% "7b"] <- "Central & Eastern Europe"
finalData$Region[finalData$Region %in% "7c"] <- "Russia, Ukraine & Belarus"

model_transform <- lm (HPI ~ factor(Region) + Internet + LifeExpect + log(GovHealth) + log(PriHealth) + PPP  + log(PopDensity),
                       data = finalData)

anova(model_transform)
summary(model_transform)
plot(model_transform)
```

## Residual Plots for final model - should we transform PPP by taking its log? Lower p-value but better fit with residual

```{r}
res <- model_transform$residuals
plot(model_transform$fitted.values, finalData$HPI)
plot(model_transform)
ggplot(data = HPIdata, aes(x = log(PPP), y = HPI)) + geom_point()  +
  aes(colour = factor(RegionNum)) + theme(legend.position = "right") + labs(title = "Purchasing Power")
finalWithRes <- cbind.data.frame(finalData, res)
In <- ggplot(data=finalWithRes, aes(x = Internet, y = res)) + geom_point() + geom_hline(yintercept = 0) + labs(title = "Internet")
Li <- ggplot(data=finalWithRes, aes(x = LifeExpect, y = res)) + geom_point() + geom_hline(yintercept = 0) + labs(title = "Life Expectency")
Go <- ggplot(data=finalWithRes, aes(x = log(GovHealth), y = res)) + geom_point() + geom_hline(yintercept = 0) + labs(title = "GovHealth")
Pri <- ggplot(data=finalWithRes, aes(x = log(PriHealth), y = res)) + geom_point() + geom_hline(yintercept = 0) + labs(title = "PriHealth")
PPP <- ggplot(data=finalWithRes, aes(x = PPP, y = res)) + geom_point() + geom_hline(yintercept = 0) + labs(title = "Purchasing Power")
Pop <- ggplot(data=finalWithRes, aes(x = log(PopDensity), y = res)) + geom_point() + geom_hline(yintercept = 0) + labs(title = "PopDensity")
grid.arrange(In, Li, Go, ncol = 2)
grid.arrange(Pri, PPP, Pop, ncol = 2)
#PPP
```


## Interaction plot of the final model

```{r}
# Interaction with region & internet: 3b (Middle East / South West Asia: 0.02561), 4a (Southern & Central Africa: 0.02042), 7b (Central & Eastern Europe: 0.00182)
# Anova: 0.11500
region_internet <- lm (HPI ~ factor(Region) + Internet + LifeExpect + log(GovHealth) + log(PriHealth) + PPP  + log(PopDensity) + factor(Region)*Internet, data = finalData)

# Interaction with region & LifeExpect: 3a (North Africa: 0.006482), 3b (Middle East / South West Asia: 1.57e-07), 4a (Southern & Central Africa: 0.000214), 4b(East Africa: 1.04e-06), 5a(South Asia: 0.011167), 7b (Central & Eastern Europe: 0.001128)
# Anova: 2.579e-05
region_le <- lm (HPI ~ factor(Region) + Internet + LifeExpect + log(GovHealth) + log(PriHealth) + PPP  + log(PopDensity) + factor(Region)*LifeExpect, data = finalData)

# Interaction with region & log(GovHealth): 3a (North Africa: 0.011573), 4a (Southern & Central Africa: 0.000588), 4b (East Africa: 0.022740), 7a (Central Asia & Caucuses: 0.000625), 7b (Central & Eastern Europe: 0.004876)
# Anova: 0.0009153
region_gh <- lm (HPI ~ factor(Region) + Internet + LifeExpect + log(GovHealth) + log(PriHealth) + PPP  + log(PopDensity) + factor(Region)*log(GovHealth), data = finalData)

# Interaction with region & log(PriHealth): 2b (North America: 0.02771), 4a (Southern & Central Africa: 0.00378), 5a (South Asia: 0.03188), 6b (Wealthy East Asia: 0.04037), 7a (Central Asia & Caucuses: 0.01119)
# Anova: 0.001166 
region_ph <- lm (HPI ~ factor(Region) + Internet + LifeExpect + log(GovHealth) + log(PriHealth) + PPP  + log(PopDensity) + factor(Region)*log(PriHealth), data = finalData)

# Interaction with region & PPP: 4a (Southern & Central Africa: 0.022310), 4c (West Africa: 0.003436), 7a (Central Asia & Caucuses: 0.037789)
# Anova: 0.004437
region_ppp <- lm (HPI ~ factor(Region) + Internet + LifeExpect + log(GovHealth) + log(PriHealth) + PPP  + log(PopDensity) + factor(Region)*PPP, data = finalData)

#Interaction with region and log(PopDensity): 2b (North America: 0.002350), 2d (Nordic Europe: 0.001224), 3b (Middle East / South West Asia: < 2e-16), 4b (East Africa: 2.35e-07), 4c(West Africa: 8.28e-06), 5a (South Asia: 1.35e-05), 7b (Central & Eastern Europe: 0.024437), 7c (Russia, Ukraine & Belarus: 0.045916)
# Anova: < 2.2e-16
region_pop <- lm (HPI ~ factor(Region) + Internet + LifeExpect + log(GovHealth) + log(PriHealth) + PPP  + log(PopDensity) + factor(Region)*log(PopDensity), data = finalData)

# Interaction with log(PriHealth) & log(GovHealth)
# Anova: 0.93443
gov_pri <- lm (HPI ~ factor(Region) + Internet + LifeExpect + log(GovHealth) + log(PriHealth) + PPP  + log(PopDensity) + log(GovHealth)*log(PriHealth), data = finalData)

anova(region_internet)
summary(region_ph)
plot(gov_pri)
```

## Extra Sum of Squares for WomenParliamant and LaborForceFemale
 - Need to see whether we should delete Gov Health from these tests
```{r}
# With WomenParliament 
model_wom_parl <- lm (HPI ~ factor(Region) + Internet + LifeExpect +  log(PriHealth) + log(GovHealth) + PPP +
                         log(PopDensity)+ WomenParliamant,
                       data = finalData)
summary(model_wom_parl) #R2 of 0.7832
# With Labor Force Female 
model_lab_force <- lm (HPI ~ factor(Region) + Internet + LifeExpect + log(PriHealth)+ log(GovHealth) + PPP +
                         log(PopDensity)+ LaborForceFemale*LaborForceFemale,
                       data = finalData)
summary(model_lab_force) #0.7908
anova(model_transform, model_lab_force)
anova(model_transform, model_wom_parl)
anova(model_transform, model_lab_force) # p-val of 0.0002039 ***
anova(model_transform, model_wom_parl)
```

## HPI map

```{r}
library(maps) 
# load world map
BaseMap <- map_data("world")
colnames(BaseMap) <- c(colnames(BaseMap)[1:4], "Country", "subregion")
BaseMap$Country <- as.character(BaseMap$Country)

# Fix BaseMap Country names to match with HPI data
BaseMap$Country[BaseMap$Country == "Republic of Congo"] <- "Congo, Rep."
BaseMap$Country[BaseMap$Country == "Democratic Republic of the Congo"] <- "Congo, Dem. Rep."
BaseMap$Country[BaseMap$Country == "Hong Kong"] <- "Hong Kong"
BaseMap$Country[BaseMap$Country == "South Korea"] <- "Korea, Rep."
BaseMap$Country[BaseMap$Country == "Burma"] <- "Trinidad"
BaseMap$Country[BaseMap$Country == "UK"] <- "United Kingdom"
BaseMap$Country[BaseMap$Country == "USA"] <- "United States"

# Join BaseMap and HPIdata
BaseMap$Country <- as.character(BaseMap$Country)
HPIdata$Country <- as.character(HPIdata$Country)
HPIwithLocation <- inner_join(BaseMap, HPIdata, by = "Country")

p09 <- ggplot() + geom_polygon(data=filter(HPIwithLocation, Year == 2009), aes(x=long, y=lat, group=group, fill = HPI), 
                             color="white", size = 0.2) + 
  scale_fill_gradient(name="HPI score", 
            low = "red", high = "blue", limits = c(0,80), na.value = "grey50") +
  labs(title="HPI score in 2009")
p12 <- ggplot() + geom_polygon(data=filter(HPIwithLocation, Year == 2012), aes(x=long, y=lat, group=group, fill = HPI), 
                             color="white", size = 0.2) + 
 scale_fill_gradient(name="HPI score", 
            low = "red", high = "blue", limits = c(0,80), na.value = "grey50") +
  labs(title="HPI score in 2012")
p16 <- ggplot() + geom_polygon(data=filter(HPIwithLocation, Year == 2016), aes(x=long, y=lat, group=group, fill = HPI), 
                             color="white", size = 0.2) + 
  scale_fill_gradient(name="HPI score", 
            low = "red", high = "blue", limits = c(0,80), na.value = "grey50") +
  labs(title="HPI score in 2016")

p09
p12
p16
# visual with HPI rank
# Plot residuals to see what we predicted was different from original data
```

## Plot showing multicollinearity between Private Health and Gov Health 
```{r}
ggplot(data = finalData, aes(x = GovHealth, y = PriHealth)) + geom_point()  + labs(title = "Multicollinearity between Private and Government Spending on Healthcare")
```
